Algorithm explanation

### 2013-08-25 article/algo.tex
Explanation for gen_constr

$C_v$ is defined by well-founded induction on $v$ with respect to the
relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
\begin{itemize}
\item For $v \in V_\Rightarrow$;
\begin{align*}
C_v =
 \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
 \bigcup_{0 < i \leq K}
 \left\lbrace
  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
 \right\rbrace \\
 & \hspace{2cm} \cup
 \left\lbrace
  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
 \right\rbrace \cup
 \bigcup_{k_{succ(v)} < i \leq K}
 \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
\end{align*}
\item For $v \in V_T$ such that
  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
  where $\mathbf{a}_0$ and $b_0$ are constants;
\begin{align*}
C_v =
\left\lbrace
 \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
 b_v = \lambda_v b_0
\right\rbrace
\end{align*}
\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
\begin{align*}
C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
\right)
\end{align*}
\item For $v_\bot$;
\begin{align*}
C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
\end{align*}
\end{itemize}

@@

\paragraph{Conjunction split}
For the first case previously mentioned, the unsatisfiable core
$\mathcal{U}$ contains coeffcient constraint expressions with a
coefficient parameter of certain predicate variable.  If the predicate
variable $P$ is the cause of unsatisfiability, some constraint
expressions in the unsatisfiable core have a parameter $a_{P,i}$
and/or $b_P$ in the following way.
\begin{align*}
& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
& \left( \exists i, j, k;
\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
\right\rbrace \subseteq \mathcal{U} \right)
\end{align*}

@@

\paragraph{Disjunction split}
For the second case, the unsatisfiable core $\mathcal{U}$ contains
coefficient constraints that unifies induced linear expressions from
each term in disjunctions to be the same expression.  Similarly as
previous, if the predicate variable $P$ is the cause of
unsatisfiability, some constraint expressions in the unsatisfiable
core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
\begin{align*}
& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
& \left( \exists i, j, k;
\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
\right\rbrace \subseteq \mathcal{U} \right)
\end{align*}

@@

@@

\paragraph{Algorithm}
To avoid this, our algorithm tries to unify a disjunctions of solution
and determine the necessity of split in conjunctions.

The constraints are first built the weakest so that all conjunctions
are not yet unified.

The strength matters that the
predicate variables are used as the same expressions depending on
different appearances of the predicate variable.

The strngeth of the constraint is controlled by a set of vertices $V$
in the pseudo-code.  Predicate variable vertices in $V$ are treated
strictly unique across the different appearances of the predicate
variables in Horn clauses.

$\genConstr$ procedure produces the constraint for a given vertex $v$.

\begin{algorithm}
\caption{$ \genConstr (G, v, visited) $}\label{alg:genConstr}
\begin{algorithmic}
\STATE {$C \gets \emptyset$}
\FORALL {$v_\Rightarrow \in preds_G(v)$}
  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
    \IF {$\varphi (v_p)$ is a predicate variable}
      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
    \ELSE
      \STATE { $id$ }
    \ENDIF
  \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}

>>>>

$C_v$ is defined by well-founded induction on $v$ with respect to the
relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
\begin{itemize}
\item For $v \in V_\Rightarrow$;
\begin{align*}
C_v & =
 \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
 \bigcup_{0 < i \leq K}
 \left\lbrace
  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
 \right\rbrace \\
 & \hspace{2cm} \cup
 \left\lbrace
  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
 \right\rbrace \cup
 \bigcup_{k_{succ(v)} < i \leq K}
 \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
\end{align*}
\item For $v \in V_T$ such that
  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
  where $\mathbf{a}_0$ and $b_0$ are constants;
\begin{align*}
C_v =
\left\lbrace
 \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
 b_v = \lambda_v b_0
\right\rbrace
\end{align*}
\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
\begin{align*}
C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
\right)
\end{align*}
\item For $v_\bot$;
\begin{align*}
C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
\end{align*}
\end{itemize}

@@

\paragraph{Conjunction split}
For the first case previously mentioned, the unsatisfiable core
$\mathcal{U}$ contains coeffcient constraint expressions with a
coefficient parameter of certain predicate variable.  If the predicate
variable $P$ is the cause of unsatisfiability, some constraint
expressions in the unsatisfiable core have a parameter $a_{P,i}$
and/or $b_P$ in the following way.
\begin{align*}
% TODO:
& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
& \left( \exists i, j, k;
\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
\right\rbrace \subseteq \mathcal{U} \right)
\end{align*}

@@

\paragraph{Disjunction split}
For the second case, the unsatisfiable core $\mathcal{U}$ contains
coefficient constraints that unifies induced linear expressions from
each term in disjunctions to be the same expression.  Similarly as
previous, if the predicate variable $P$ is the cause of
unsatisfiability, some constraint expressions in the unsatisfiable
core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
\begin{align*}
% TODO:
& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
& \left( \exists i, j, k;
\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
\right\rbrace \subseteq \mathcal{U} \right)
\end{align*}

@@

Consider these two ways of transformation, we show a pseudo-code for
\transGraph~procedure.  The procedure is given the graph $G$, the
predicate templateset $T$, and the unsatisfiable core
$\mathcal{U}$. It returns the transformed graph $G'$ and the new
templateset $T'$.

\begin{algorithm}
\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
\begin{algorithmic}
% TODO:
\RETURN {$\left( G', T' \right)$}
\end{algorithmic}
\end{algorithm}

@@

\paragraph{Algorithm}
To avoid this, our algorithm lets a predicate variable $P$ which is
disjunctively implied on the Horn graph $G$ to have a linear
expression that is consequences of all terms in the disjunction.  This
prevents the possibility of choices for disjunctions to be propagated
to succeeding vertices. The simple linear expression is propagated
instead.

On the other hand, the algorithm allows a predicate variable $P$ that
appears on multiple implications or appears multiple times in a single
implication to take different linear expressions as a premise at
different appearances.  This is for the algorithm to determine the
cause of unsatisfiability efficiently.  It is done by copying a
predicate templates and its parameters' constraint with appropriate
renaming.

The algorithm first execute this duplication on all predicate
variables with multiple appearances.  It is the weakest constraint
that the algorithm makes for the certain $G$.  Then, the algorithm
gradually reduces the number of vertices that are duplicated by each
time testing the satisfiability of the constraint.  When the
constraint bebcome unsatisfiable at the certain iteration, the
algorithm can determine a predicate vertex which has to be split in a
conjunctive template.  If all the predicate variable constraints are
no longer duplicated and the constraint is the strongest, any model
for such constraint gives a simple solution under the Horn graph $G$
and the templateset $T$.

Note that duplicating the constraint in a na\"{i}ve way may cause the
constraint to exponentially grow in size.  Our algorithm prevent it by
simplifying a predicate vertex's constraint $C_v$ by quantifier
elimination when $v$ is duplicated in generating linear constraint.
For a predicate variable vertex $v_P$, variables except for the
coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
are eliminated.  Due to the constraints generation process, the
simplified constraint is also linear and can be efficiently check the
satisfiability.

We finally show the modified version of pseudo-code for $\genConstr$
procedure.  It produces the constraint $C_v$ for a given vertex $v$,
considering constraint duplication and simplification.

\begin{algorithm}
\caption{$ \genConstr (G, v, V_\star) $}\label{alg:genConstr}
\begin{algorithmic}
\STATE {$C \gets \emptyset$}
\FORALL {$v_\Rightarrow \in preds_G(v)$}
  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
    \IF {$\varphi (v_p)$ is a predicate variable}
      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
    \ELSE
      \STATE { $id$ }
    \ENDIF
  \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}
